---
title: "Formative Assessment 5"
author: "Vera Aguila"
date: "2024-09-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 8.18

List all samples of size *n*=2 that are possible (with replacement) from the population in Problem 8.17. 

The population consists of the credit hours X={9,12,15}, and each has a probability p(x)=1/3.

Since we are sampling with replacement, the possible pairs of 𝑋 are: 

{(9,9),(9,12),(9,15),(12,9),(12,12),(12,15),(15,9),(15,12),(15,15)}


Use R to plot the sampling distribution of the mean to show that $\mu_{\bar{x}} = \mu$,and show that $\sigma_{\bar{x}}^2 = \sigma^2 / 2$.

```{r eight}
X <- c(9, 12, 15)
p <- rep(1/3, 3) 

samples <- expand.grid(X1=X, X2=X)

samplemeans <- rowMeans(samples)

sampling_table <- data.frame(Sample_1 = samples$X1, Sample_2 = samples$X2, Sample_Mean = samplemeans)

print(sampling_table)

meansamplemean <- mean(samplemeans)

populationmean <- sum(X * p)

varsamplemean <- var(samplemeans)

populationvariance <- sum((X^2 * p)) - populationmean^2

expectedvarsamplemean <- populationvariance / 2

list(
  samples = samples,
  samplemeans = samplemeans,
  meansamplemean = meansamplemean,
  populationmean = populationmean,
  varsamplemean = varsamplemean,
  expectedvarsamplemean = expectedvarsamplemean
)

```



## Problem 8.21

A population consists of the four numbers 3, 7, 11, and 15. Consider all possible samples of size 2 that can be drawn with replacement from this population. Find:

(a) the population mean

```{r, include=FALSE}
population <- c(3, 7, 11, 15)
popmean <- mean(population)
popmean

```

(b) the population standard deviation

```{r, include=FALSE}
popsd <- sd(population)
popsd

```

(c) the mean of the sampling distribution of means

```{r}
samples <- expand.grid(population, population)
samplemeans <- rowMeans(samples)
samplingdistmean <- mean(samplemeans)
samplingdistmean

```

(d) the standard deviation of the sampling distribution of means.

```{r}
samplingdistsd <- sd(samplemeans)
samplingdistsd

```

Verify parts (c) and (d ) directly from (a) and (b) by using suitable formulas. Using the formulas:

-   Mean of the sampling distribution = Population mean
-   Standard deviation of the sampling distribution = Population standard deviation divided by the square root of the sample size

```{r}
verifysd <- popsd / sqrt(2)
verifysd

```

## Problem 8.34

Find the probability that of the next 200 children born. Assume equal probabilities for the births of boys and girls.

Given: \* The probability of a boy being born, p = 0.5 \* The number of trials, n = 200

```{r, include=FALSE}
n <- 200          
b <- 0.5          
g <- 1 - b
meanprop <- b 
seprop <- sqrt(b * g / n)
```

(a) less than 40% will be boys

-   40% of 200 children = 0.40×200=80 boys.
-   We want to find P(X\<80).

```{r}
phata <- 0.40  
za <- (phata - meanprop) / seprop  
za


proba <- pnorm(za)  
proba

```

(b) between 43% and 57% will be girls

-   43% and 57% of 200 children correspond to: 0.43×200=86 girls, or 114 boys.
-   57% of 200 = 0.57×200=114 girls, or 86 boys.
-   We want to find P(86≤X≤114), where X is the number of girls.

```{r}
phatb1 <- 1 - 0.57  
phatb2 <- 1 - 0.43  

zb1 <- (phatb1 - meanprop) / seprop
zb2 <- (phatb2 - meanprop) / seprop


probb <- pnorm(zb2) - pnorm(zb1)
probb


```

(c) more than 54% will be boys

-   54% of 200 children = 0.54×200=108 boys.
-   We want to find P(X\>108).

```{r}
phatc <- 0.54
zc <- (phatc - meanprop) / seprop
zc

probc <- 1 - pnorm(zc)
probc


```

## Problem 8.49

The credit hour distribution at Metropolitan Technological College is as follows:

Find $\mu$ and $\sigma^2$. Give the 25 (with replacement) possible samples of size 2, their means, and their probabilities.

**Mean:**

```{r}
x <- c(6, 9, 12, 15, 18)
px <- c(0.1, 0.2, 0.4, 0.2, 0.1)  

mu <- sum(x * px)
mu  
```

$\sigma^2$**:**

```{r}
sigma2 <- sum((x - mu)^2 * px)
sigma2

samples <- expand.grid(x, x) 
samplemeans <- rowMeans(samples)

sampleprobs <- outer(px, px, "*")  
sampleprobs <- as.vector(sampleprobs)

results <- data.frame(Sample1 = samples[,1], Sample2 = samples[,2], 
                      Mean = samplemeans, Probability = sampleprobs)


results

```
